---
title: "Multifamily Permits"
author: "Jesse Wade"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Read in Data from census

The data that I am using comes directly from the US Census Bureau. The number of multifamily permits is based on the number of Units that have been permited for more non single-family structures. This will be the sum of the the columns named

```{r Census_data}
header <- scan("https://www2.census.gov/econ/bps/County/co2301c.txt",nlines= 1, what=character(),sep=",")
header2 <- scan("https://www2.census.gov/econ/bps/County/co2301c.txt",nlines= 1,skip=1 ,what=character(),sep=",")
data <- read.csv("https://www2.census.gov/econ/bps/County/co2301c.txt",skip= 3,header = FALSE)
names(data) <- paste0(header,header2)

# Cut off the end to use only reported data
data <- data[,1:18]
```

## Remove single family data and format multifamily data correctly

```{r}
data %>% 
  select(c(1:6,10:18)) %>% 
  mutate(total_mf_bldgs=.[[7]]+.[[10]]+.[[13]],total_mf_units=.[[8]]+.[[11]]+.[[14]],total_mf_value=.[[9]]+.[[12]]+.[[15]] ) %>% 
  filter(total_mf_units != 0)-> data_1
```

Data looks ready! :) The counties that did not have any multifamily permits have been removed from the data.

### County Level Analysis 

#### Average Multifamily Permits 

A pretty low average across the U.S. at only 81.376 in the month of January 2023. Very strong left skew towards the bottom

```{r Focus_on_Top10_Counties}
data_1 %>% 
  summarise(AvgPermitsMF=mean(total_mf_units),MeidanPermitsMF=median(total_mf_units))
```
```{r Top_Counties , echo=FALSE}
data_1 %>% 
  select(CountyName,total_mf_units) %>% 
  arrange(desc(total_mf_units)) %>% 
  top_n(20,total_mf_units)-> County_top_20
knitr::kable(
 County_top_20,caption = "Top 20 Multifamily Permits in January 2023 by number of Units, County Level"
)
```


### State Level Count 

This data set is at the county level. It contains the current level of permits for the particular month by county

```{r States_Read}
data_1 %>% 
  group_by(FIPSState) %>% 
  summarise(MultifamilyPermits = sum(total_mf_units)) %>% 
  arrange(desc(MultifamilyPermits))-> State_Counts
knitr::kable(
  State_Counts %>% 
    top_n(10,MultifamilyPermits),caption = "Top 10 Multifamily Permits in January 2023, Statewise"
)
```

Only have FIPS codes, will convert over later. 
