---
title: "Single-family Permits"
author: "Jesse Wade"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(shiny)
```

# Read in Data from census

I am reading in the data from the census. It gives the count for the the year and data. This particular dataset is set for January 2023, the most recent release of data. The next permits release is on March 16, which will be February permit data at the county level.

```{r Census_data}
header <- scan("https://www2.census.gov/econ/bps/County/co2301c.txt",nlines= 1, what=character(),sep=",")
header2 <- scan("https://www2.census.gov/econ/bps/County/co2301c.txt",nlines= 1,skip=1 ,what=character(),sep=",")
data <- read.csv("https://www2.census.gov/econ/bps/County/co2301c.txt",skip= 3,header = FALSE)
names(data) <- paste0(header,header2,sep = "")

data <- data[,1:18]
```

## Remove all multifamily permit data

```{r}
data %>% 
  select(c(1:9)) ->data_1

head(data_1)
```

Data looks ready! :)

### County Level Analysis 

#### Top Countys by Permits level 

```{r Top_10_Permits}
data_1 %>% 
  select(CountyName,Permits=8) %>% 
  arrange(desc(Permits)) %>% 
  top_n(20,Permits) -> Top_sf_Permit_Counties
knitr::kable(
  Top_sf_Permit_Counties,caption = "Top 20 Single Family Permits in January 2023 by County"
)

```


### State Level Count 

This data set is at the county level. It contains the current level of permits for the particular month by county

```{r States_Read}
data %>% 
  group_by(FIPSState) %>% 
  summarise(SingleFamilyPermits = sum(Bldgs)) %>% 
  arrange(desc(SingleFamilyPermits))-> State_Counts
knitr::kable(
  State_Counts %>% 
    top_n(10,SingleFamilyPermits),caption = "Top 10 Single Family Permits in January 2023, Statewise"
)
```

Only have FIPS codes, will convert over later. 


### Adding in R-Shiny application to site. Permits by county or MSA

```{r ShinyPermitsApp, echo=FALSE }
shinyAppFile("PermitsApp.R")
```

